An Automatic Verification Framework for Software Configurations

### 2016-05-09 OSDI/intro.tex
config: fixed typo in intro

While many efforts have been proposed 
to check, troubleshoot, diagnose and repair configuration errors%
~\cite{attariyan10automating,
su07autobash, whitaker04configuration},
those tools mainly try to understand {\emph{what}} caused the error ---
they are still not on a level of
automatic verification tools used for regular program 
verification~\cite{Leino10Dafny, PiskacWZ14, BobotFMP15} that can
detect errors without executing the code.
Two main obstacles why we cannot simply apply the existing automatic 
tools and techniques to verification of configuration files are: 1) a lack
of a specification which would describe properties of configuration files, and 2) a program structure of configuration files -- they
are mainly a sequence of entries assigning some value to system variables. The language in which configuration files are written does 
not adhere to a specific grammar or syntax. In particular, the
entries in configuration files are untyped. Moreover, there are surprisingly few rules specifying constraints on entries and there
is no explicit structure policy for the entries.
Nevertheless, automated verification of configuration 
files, that could detect would be highly
desirable~\cite{wang04automatic, zhang14encore, xu15systems}.

To overcome these obstacles, researchers have proposed approaches based on
a statistical analysis and learning~\cite{wang04automatic, zhang14encore, yuan11context}
that try to infer the rules and policies how configuration files are constructed.
Rather than explicitly specifying entries' types or rules, these research directions are focused
on learning policies from a sample data set.
The learned rules are usually constructed along the following pattern: for every entry in a configuration file, 
we check if it deviates from a ``typical'' value, \ie a value computed from a large training set consisting of configuration files.
If the entry is significantly different that the typical value, we suspect that it could be a potential configuration error.
The downside of this approach is that these learning efforts are limited to simplistic 
configuration errors, such as type errors and syntax errors, or they 
heavily rely on template-based inference~\cite{zhang14encore}. As we pointed out earlier, since configuration files are not written in a typed language with a fixed syntax,
type errors and syntax errors are relatively rare and easy to detect. However, many sophisticated configuration errors, 
which are difficult to template in practice, cannot be detected this way. As an illustration of such an error consider the ordering errors, which
happen when some entries are inserted in a wrong order. 
For example, if in a PHP configuration file an entry {\tt extension = mysql.so} appears 
before {\tt extension = recode.so},  
it would lead to a crash error, where the Apache server cannot start due to the segmentation fault error. The correct ordering 
should be {\tt extension = recode.so} before 
{\tt extension = mysql.so}~\cite{yin11anempirical}.
These types of errors cannot be detected by existing
learning efforts, since it is hard to build a corresponding template ~\cite{xu15systems}: we would need to capture ordering relations of all the entries. Some of 
them can appear in an arbitrary ordering and the resulting templates would be too large and not efficient for a practical use.

@@

In this paper we present \app, the first automatic verification framework
for general software configurations. \app is based on a collection of powerful learning 
algorithms, that do not necessarily depend on templates. The learning process takes as input a
 large sample of configuration files. The process is language-agnostic and it works for any type of configuration files, 
but all of the files in the sample need to be of the same type (such as  MySQL or httpd configuration files). From that sample
\app learns an abundant set of rules specifying various properties that hold on the given sample. The files in the sample might contain
errors -- we are, therefore, using probabilistic learning to derive a set of accurate rules. One can see this learning process as a
way of deriving  
a specification for configuration files. The learning process works in two phases. In the first phase \app analyzes the given dataset and generates a 
well-structured and probabilisticly-typed 
intermediate representation. In the second phase \app derives rules and constraints by analyzing
the intermediate representations.  The set of learned rules is accurate and we are able to efficiently check 
the correctness of the configuration files
of interest and detect potential errors.
Compared with previous efforts,
\app does not necessarily rely on users' templates 
and is capable of detecting intricate configuration errors that
cannot be identified by existing work (details in $\S$\ref{sec-motiv}).

Building such an automatic verification framework for
configuration files, nevertheless, requires addressing several challenges. 
First, for variables of one types we learn different rules than for variables of different type. For example for integer variables $x$ and $Y$ we can learn that $X \leq Y$, while 
for strings we might want to learn a substring relation. We, therefore, first need to assign a type to every variable in a configuration file during the transformation phase.
However,
the type of a variable cannot always be fully determined 
from a single value. 
For example, an entry {\tt temp\_dir = 300} assigns to variable {\tt temp\_dir} either a directory named ``300'' or sets the size of that directory to 300 (an integer).
Some of existing type inference 
work would report this is an error, because {\tt temp\_dir} should be assigned
an integer~\cite{zhang14encore}. We address this problem by introducing the concept of
{\emph{probabilistic types}}.
Rather than assigning only one variable to a single type, 
we assign several types with their probability distributions. 
The entry in the above example might be assigned 
the following probabilistic type
{\tt \{temp\_dir, 300, [(File, 60\%), (Int, 40\%)]\}}.
Using the probabilistic types model,
we can generate a more accurate language model,
thus significantly improving our verification capabilities.

>>>>

While many efforts have been proposed 
to check, troubleshoot, diagnose and repair configuration errors%
~\cite{attariyan10automating,
su07autobash, whitaker04configuration},
those tools mainly try to understand {\emph{what}} caused the 
error -- they are still not on a level of
automatic verification tools used for regular program 
verification~\cite{Leino10Dafny, PiskacWZ14, BobotFMP15} that can
detect errors without executing the code.
Two main obstacles why we cannot simply apply the existing automatic 
tools and techniques to verification of configuration files are: 1) a lack
of a specification which would describe properties of configuration files, and 2) a program structure of configuration files -- they
are mainly a sequence of entries assigning some value to system variables. The language in which configuration files are written does 
not adhere to a specific grammar or syntax. In particular, the
entries in configuration files are untyped. Moreover, there are surprisingly few rules specifying constraints on entries and there
is no explicit structure policy for the entries.
Thus, automated verification of configuration 
files would be highly
desirable~\cite{wang04automatic, zhang14encore, xu15systems}.

To overcome these obstacles, researchers have proposed approaches based on
a statistical analysis and learning%
~\cite{wang04automatic, zhang14encore, yuan11context}
that try to infer the rules and policies 
how configuration files are constructed.
Rather than explicitly specifying entries' types or rules, 
these efforts are focused on learning policies from a sample dataset.
The learned rules are usually constructed along the following way: 
for every entry in a configuration file, 
they check if it deviates from a ``typical'' value, \ie, 
a value computed from a large training set 
consisting of configuration files.
If the entry is significantly different that the typical value, 
they suspect that it could be a potential configuration error.
The downside of these approaches is that these learning approaches 
are limited to simplistic configuration errors, 
such as type errors and syntax errors, or they 
heavily rely on template-based inference~\cite{zhang14encore}. 
As we pointed out earlier, 
since configuration files are not written in a typed language 
with a fixed syntax,
type errors and syntax errors are relatively rare and easy to detect. 
However, many sophisticated configuration errors, 
which are difficult to template in practice, 
cannot be detected by this way. 
As an illustration of such an error consider the ordering errors, 
which happens when some entries are inserted in a wrong order. 
For example, if in a PHP configuration file 
an entry {\tt extension = mysql.so} appears 
before {\tt extension = recode.so},  
it would lead to a crash error, where the Apache server cannot start 
due to the segmentation fault error. The correct ordering 
should be {\tt extension = recode.so} before 
{\tt extension = mysql.so}~\cite{yin11anempirical}.
These types of errors cannot be detected by existing
learning efforts, 
since it is hard to build a corresponding template~\cite{xu15systems}: 
we would need to capture ordering relations of all the entries. 
Some of them can appear in an arbitrary ordering, 
and the resulting templates would be too large 
and not efficient for a practical use.

@@

In this paper, we present \app, the first automatic verification framework
for general software configurations. \app is based on a collection of powerful learning 
algorithms, that do not necessarily depend on templates. The learning process takes as input a
 large sample of configuration files. The process is language-agnostic and it works for any type of configuration files, 
but all of the files in the sample need to be of the same type (such as
MySQL or httpd configuration files). From that sample,
\app learns an abundant set of rules specifying various properties that hold on the given sample. The files in the sample might contain
errors -- we are, therefore, using probabilistic learning to derive a set of accurate rules. One can see this learning process as a
way of deriving  
a specification for configuration files. 
The learning process works in two phases. In the first phase, 
\app analyzes the sample dataset and generates a 
well-structured and probabilistically-typed 
intermediate representation. 
In the second phase, \app derives rules and constraints by analyzing
the intermediate representations.  The set of learned rules is accurate and we are able to efficiently check 
the correctness of the configuration files
of interest and detect potential errors.
Compared with previous efforts,
\app does not necessarily rely on users' templates 
and is capable of detecting intricate configuration errors that
cannot be identified by existing work (details in $\S$\ref{sec-motiv}).

Building such an automatic verification framework for
configuration files, nevertheless, requires addressing several challenges. 
First, for variables of one type 
we learn different rules than for variables of different types. 
For example, for integer variables $X$ and $Y$, 
we can learn that $X \leq Y$, while 
for strings we might want to learn a substring relation. We, therefore, first need to assign a type to every variable in a configuration file during the transformation phase.
However,
the type of a variable cannot always be fully determined 
from a single value. 
For example, an entry {\tt temp\_dir = 300} assigns to variable {\tt temp\_dir} either a directory named ``300'' or sets the size of that directory to 300 (an integer).
Some of existing type inference 
work would report this is an error, because {\tt temp\_dir} should be assigned
an integer~\cite{zhang14encore}. We address this problem by introducing the concept of
{\emph{probabilistic types}}.
Rather than assigning only one variable to a single type, 
we assign several types with their probability distributions. 
The entry in the above example might be assigned 
the following probabilistic type
{\tt \{temp\_dir, 300, [(File, 60\%), (Int, 40\%)]\}}.
Using the probabilistic types model,
we can generate a more accurate language model,
thus significantly improving our verification capabilities.
